---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
first, we want to unzip and read the data into a data.frame
``` {r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```
verify, that something has been loaded:
``` {r}
names(activity)
```

change the `date` from a factor variable to a date variable:
``` {r}
activity$date <- as.Date(activity$date)
```


## What is mean total number of steps taken per day?

calculate the total number of steps per day:
``` {r}
nStepsPerDay <- aggregate(activity$steps, list(activity$date), sum)
names(nStepsPerDay) <- c("Date", "total steps")
```

here is a bar bplot for the total number of steps per day:
``` {r}
library(ggplot2)
g <- ggplot(nStepsPerDay, aes(x=Date, y = `total steps`))
g <- g + geom_bar(stat = "identity", position = "stack")
g <- g + ggtitle("number of steps per day")
g
```

and here is the histogram:

``` {r}
g <- ggplot(nStepsPerDay, aes(`total steps`))
g <- g + geom_histogram()
g <- g + ggtitle("histogram of number of steps per day")
g
```

the mean is:
``` {r}
mean(nStepsPerDay$`total steps`, na.rm = T)
```

and the median is:
``` {r}
median(nStepsPerDay$`total steps`, na.rm = T)
```


## What is the average daily activity pattern?
Because we are interested in plotting a time series, it seems like a better idea to have a rough value rather than none, so we filter out `NA` values. We get the mean number of steps per 5-minute time period. 
``` {r}
nStepsPerTimeInterval <- aggregate(activity$steps, list(activity$interval), mean, na.rm=T)
names(nStepsPerTimeInterval) <- c("interval", "average steps")
```

here the time series of mean steps per time interval:
``` {r}
g <- ggplot(nStepsPerTimeInterval, aes(x=interval, y = `average steps`))
g <- g + geom_line()
g <- g + ggtitle("average number of steps per time interval")
g
```

the interval with the most steps in average is:
``` {r}
nStepsPerTimeInterval[which.max(nStepsPerTimeInterval$`average steps`), "interval"]
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
